Synthesizing Symmetric Lenses

### 2018-07-03 papers/SymmetricLenses/main.tex
Added proof of Theorem 1

\begin{center}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{(1-\Probability_1)*\Probability_2})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
\end{center}

@@

\fi

\end{document}

>>>>

\begin{center}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)*\Probability_2}{1 - \Probability_1 * \Probability_2}})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
\end{center}

@@

\fi
\section{Forgetful Symmetric Lenses}
\begin{theorem*}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem*}
The following lemma is
\begin{proof}
\begin{enumerate}

\item
($S \; |_1 \; \varnothing \equiv^S S$)

For all $s \in \mathcal{L}(S \; | \; \varnothing)$, then $P_{S \; |_1 \; \varnothing} = 1 * P_S(s) = P_S(s)$.
\item
($S \cdot \varnothing \equiv^S \varnothing$)

For all $s \in \mathcal{L}(S \cdot \varnothing)$, the $P_{S \cdot \varnothing}(s) = \sum_{s_1 \cdot s_2}P_S(s_1) * P_{\varnothing}(s_2)$, but this sum is empty, hence $P_{S \cdot \varnothing}(s) = 0 = P_{\varnothing}(s)$.
\item
($\varnothing \cdot S \equiv^S \varnothing$)

Similar to the previous case.
\item
($(S \cdot S') \cdot S'' \equiv^S S \cdot (S' \cdot S'')$). Let $s \in \mathcal{L}(S \cdot S' \cdot S'')$

Then
\begin{align*}
P_{(S \cdot S') \cdot S''}(s) &= \sum_{s=s_4 \cdot s_3}P_{S \cdot S'}(s_4) * P_{S''}(s_3)\\
&= \sum_{s=s_4 \cdot s_3}\left(\sum_{s_4=s_1 \cdot s_2} P_{S}(s_1) * P_{S'}(s_2)\right)* P_{S''}(s_3)\\
&= \sum_{s=s_1\cdot s_2 \cdot s_3} P_{S}(s_1) * P_{S'}(s_2) * P_{S''}(s_3)\\
P_{S \cdot (S' \cdot S'')}(s) &= \sum_{s=s_1\cdot s_4} P_{S}(s) * \left(\sum_{s_4=s_2\cdot s_3}P_{S'}(s_2) * P_{S''}(s_3)\right)
\end{align*}
\item
($(S \; |_{p_1} \; S') \; |_{p_2} \; S'' \equiv^S S \; |_{p_1 * p_2} \; (S' \; |_{\frac{(1-p_1)*p_2}{1-p_1*p_2}} \; S'')$)

Let $s \in \mathcal{L}(S \; |\; S' \; | \; S'')$. Then
\begin{align*}
P_{(S \; |_{p_1} \; S') \; |_{p_2} \; S''}(s) &= p_2 * P_{S \; |_{p_1} \; S'}(s) + (1-p_2)P_{S''}(s)\\
&= (p_2 * (p_1 * P_S(s) + (1-p_1) * P_{S'}(s))) + (1-p_2)P_{S''}(s)\\
&= (p_2 * (p_1 * P_S(s) + P_{S'}(s)-p_1* P_{S'}(s))) + (1-p_2)P_{S''}(s)\\
&= p_2*p_1 * P_S(s) + p_2*P_{S'}(s)-p_2*p_1* P_{S'}(s) + (1-p_2)P_{S''}(s)\\
&= p_2*p_1 * P_S(s) + (1-p_1) * p_2 * P_{S'}(s)+ (1-p_2)P_{S''}(s)\\
S \; |_{p_1 * p_2} \; (S' \; |_{\frac{(1-p_1)*p_2}{1-p_1*p_2}} \; S'') &= (p_1p_2)  P_S(s) + (1-p_1p_2)  \left(\frac{(1-p_1)p_2}{1-p_1p_2}  P_{S'}(s) + \left(1-\frac{(1-p_1)p_2}{1-p_1p_2}\right)P_{S''}(s)\right)
\end{align*}
\item
($S \; |_p \; T \equiv^S T \; |_{1-p} \; S$)

For all $s \in \mathcal{L}(S \; | \; T)$, we have
$$P_{S \; |_p \; T}(s) = p*P_S(s) + (1-p)*P_T(s) = (1-p) * P_T(s) + (1-(1-p))*P_s(s) = P_{T \; |_{1-p} \; S}(s)$$
\item
($S \cdot (S' \; |_p \; S'' ) \equiv^S (S \cdot S') \; |_p \; (S \cdot S'')$)

For all $s \in \mathcal{L}(S \cdot (S' \sep S''))$ we have
\begin{align*}
P_{S \cdot (S' \; |_p \; S'' )} &= \sum_{s_1 \cdot s_2 = s}P_S(s_1) * P_{S' \; |_p \; S''}(s)\\
&= \sum_{s_1 \cdot s_2 = s}P_S(s_1) * (p * P_{S'}(s_2) + (1-p) * P_{S''}(s_2))\\
&= \sum_{s_1 \cdot s_2 = s}p * P_S(s_1) * P_{S'}(s_2) + \sum_{s_1 \cdot s_2 = s}(1-p) * P_S(s_1) * P_{S''}(s_2)\\
P_{(S \cdot S') \; |_p \; (S \cdot S'')}(s) &= p*P_{S \cdot S'}(s) + (1-p)*P_{S \cdot S''}(s)
\end{align*}
\item
$((S' \; |_p \; S'') \cdot S \equiv^S (S' \cdot S) \; |_p \; (S'' \cdot S))$

For all $s \in \mathcal{L}((S' \; | \; S'') \cdot S)$,  we have
\begin{align*}
P_{(S' \; |_p \; S'') \cdot S}(s) &= \sum_{s_1 \cdot s_2 = s}P_{S' \; |_p \; S''}(s_1) * P_S(s_2)\\
&= \sum_{s_1 \cdot s_2 = s}(p * P_{S'}(s_1) + (1-p) * P_{S''}(s_1))* P_S(s_2)\\
&= \sum_{s_1 \cdot s_2=s}p*P_{S'}(s_1) * P_S(s_2) + \sum_{s_1 \cdot s_2 = s}(1-p) * P_{S''}(s_1) * P_S(s_2)\\
P_{(S' \cdot S) \; |_p \; (S'' \cdot S)}&= p * P_{S' \cdot S}(s) + (1-p) * P_{S'' \cdot S}(s)
\end{align*}
\item
($\varepsilon \cdot S \equiv^S S$)

For all $s \in \mathcal{L}(S)$, we have
$$P_{\epsilon \cdot S}(s) = \sum_{s_1 \cdot s_2 = s}P_{\epsilon}(s_1) * P_S(s_2) = P_{\epsilon}(\epsilon) * P_S(s_2) = P_S(s)$$
since $s_2 = s$ in the computation above.
\item
($S \cdot \epsilon \equiv^S S$)

Similar to the previous case.
\item
($S^{*p} \equiv^S \epsilon \; |_{(1-p)} \; (S^{*p} \cdot S)$)

Let $s \in \mathcal{L}(S^*)$, in which case we have
\begin{align*}
P_{\epsilon \; |_{1-p} \; (S^{*p} \cdot S)}(s) &= (1-p) * P_{\epsilon}(s) + p * P_{S^{*p} \cdot S}(s)\\
&= (1-p) * P_{\epsilon}(s) + p * \sum_{a \cdot b = s}\left(\sum_n \sum_{s_1 \ldots \cdot s_n=a} p^n (1-p) \prod_{i=1}^n P_S(s_i)\right) * P_S(b)\\
&= (1-p) * P_{\epsilon}(s) + \sum_{a \cdot b = s}\left(\sum_n \sum_{s_1 \cdot \ldots \cdot s_n=s} p^{n+1} (1-p) \prod_{i=1}^n P_S(s_i) * P_S(b)\right)\\
&= (1-p) * P_{\epsilon}(s) + \sum_n \sum_{s_1 \cdot \ldots \cdot s_{n+1}=s} p^{n+1} (1-p) \prod_{i=1}^{n+1} P_S(s_i)\\
&= (1-p) * P_{\epsilon}(s) + \left(\sum_n \sum_{s_1 \cdot \ldots \cdot s_n = s} p^n * (1-p) * \prod_{i=1}^n P_S(s_i)\right) - (1-p) * (\mathbbm{1}_{s = \varepsilon})
\end{align*}
Since
$$
P_{S^{*p}}(s) = \sum_n \sum_{s_1 \cdot \ldots \cdot s_n = s} p^n * (1-p) * \prod_{i=1}^n P_S(s_i)
$$
then if $s = \epsilon$, the terms $(1-p) * P_{\epsilon}(s)$ and $(1-p) * \mathbbm{1}_{s=\varepsilon}$ cancel each other out. Otherwise, if $s \neq \epsilon$, then $(1-p) * P_{\epsilon}(s) = 0 = (1-p) * (\mathbbm{1}_{s = \varepsilon})$ from which the result follows.
\item
($S^{*p} \equiv^S \epsilon \; |_{(1-p)} \; (S \cdot S^{*p})$)

Similar to the previous case.
\end{enumerate}
\end{proof}
\end{document}
